---
title: "Home"
execute:
  echo: false
  warning: false
  error: false
---

```{r "imports and variables - short trips"}
library(renv)
library(dplyr)
library(purrr)
library(ggplot2)
library(plotly)
library(paletteer)
library(nationalparkcolors)
library(scales)

source("source.R")
```

```{r "Compute some dfs - short trips"}

hours <- seq(0,3,0.25)
distance <- seq(0,50,5)

costs_dfs = list()

colour_palette <- paletteer_d("nationalparkcolors::Acadia")

plan_colours <- c(
  Open = colour_palette[[1]],
  "Open Plus" = colour_palette[[2]],
  Value = colour_palette[[3]],
  "Value Plus" = colour_palette[[4]],
  "Value Extra" = colour_palette[[5]]
)

for (
  plan_name in names(
    station_based_plans
  )
) {
  if (
    !(
      plan_name %in% c(
        "flex",
        "value_extra_workday"
      )
    )
  ) {
    colour <- plan_colours[[station_based_plans[[plan_name]]$name]]

    costs_dfs[[station_based_plans[[plan_name]]$name]] <- get_costs(
      station_based_plans[[plan_name]],
      hours,
      distance
    ) |>
    mutate(
      Colour = colour
    )
  }
}

costs_dfs[["Value"]] <- costs_dfs[["Value"]] |>
  mutate(
    open_plus_cost = costs_dfs[["Open Plus"]] |>
      select(
        total_cost
      ) |>
      pull(),
    total_cost = if_else(
      total_cost > open_plus_cost,
      open_plus_cost,
      total_cost
    )
  )

max_cost <- max(
  unlist(
    lapply(
      costs_dfs,
      \(x) max(
        x$total_cost
      )
    )
  )
)

best_cost_df <- costs_dfs |>
  bind_rows() |>
  group_by(
    hours,
    distance
  ) |>
  arrange(
    total_cost,
    plan_name,
    .by_group = TRUE
  ) |>
  mutate(
    tooltip = paste0(
      hours,
      " hours, ",
      distance,
      "km:\n---\n",
      plan_name,
      ": $",
      total_cost,
      " ($",
      time_cost,
      " + $",
      distance_cost,
      ")\n",
      collapse=""
    )
  ) |>
  slice_min(
    total_cost
  ) |>
  ungroup()
```

Fees not included:

- An additional charge of $10 per extra day applies to all trips longer than 28 days (applies to all plans)
- Surcharge on displayed prices (time + km) of 15% for minivans and 10% for family vehicles.
- Damage protection plan (applies to all plans)
- Monthly membership fee

For FLEX:

- Not considering FLEX pass
- A minimum of 4 hours will be charged even for shorter trips if the final applied cost is that of the station-based vehicles of your plan

## Best plan

```{r "best plan heatmap"}
make_best_plan_heatmap(
  best_cost_df,
  plan_colours
)
```

## Cost per trip duration

[Ottawa traffic report: average speeds (km/h) (2025)](https://www.tomtom.com/traffic-index/city/ottawa/)

|                   | City | Metro
| ---               | ---  | ---
| All time          | 34.4 | 38.8
| Morning rush hour | 31.1 | 35.0
| Evening rush hour | 24.5 | 30.0

```{r "cost per trip duration per plan"}
hours <- seq(0,3,0.25)
distance <- seq(0,350,0.1)

costs_dfs = list()

for (
  plan_name in names(
    station_based_plans
  )
) {
  if (
    !(
      plan_name %in% c(
        "flex",
        "value_extra_workday"
      )
    )
  ) {
    colour <- plan_colours[[station_based_plans[[plan_name]]$name]]

    costs_dfs[[station_based_plans[[plan_name]]$name]] <- get_costs(
      station_based_plans[[plan_name]],
      hours,
      distance
    ) |>
    mutate(
      Colour = colour
    )
  }
}

costs_dfs[["Value"]] <- costs_dfs[["Value"]] |>
  mutate(
    open_plus_cost = costs_dfs[["Open Plus"]] |>
      select(
        total_cost
      ) |>
      pull(),
    total_cost = if_else(
      total_cost > open_plus_cost,
      open_plus_cost,
      total_cost
    )
  )
```

:::{.panel-tabset}

## City, All time

```{r "cost per trip duration - City, All time"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  34.4,
  plan_colours
)
```

## Metro, All time

```{r "cost per trip duration - Metro, All time"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  38.8,
  plan_colours
)
```

## City, Morning

```{r "cost per trip duration - City, Morning"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  31.1,
  plan_colours
)
```

## Metro, Morning

```{r "cost per trip duration - Metro, Morning"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  35.0,
  plan_colours
)
```

## City, Evening

```{r "cost per trip duration - City, Evening"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  24.5,
  plan_colours
)
```

## Metro, Evening

```{r "cost per trip duration - Metro, Evening"}
make_cost_per_plan_line_chart(
  hours,
  costs_dfs,
  30,
  plan_colours
)
```

:::

