---
title: "Home"
execute:
  echo: false
  warning: false
  error: false
---

```{r "imports and variables"}
library(renv)
library(dplyr)
library(purrr)
library(ggplot2)
library(plotly)
library(paletteer)
library(nationalparkcolors)

station_based_plans = list(
  open = list(
    name = "Open",
    hourly_rate = 14.25,
    first_daily_max = 55,
    additional_day = 50,
    km_interval_1 = 75,
    km_interval_rate_1 = 0,
    km_interval_rate_2 = 0.31,
    monthly_fee = 0,
    weekend_hourly_surcharge = 0.35,
    weekend_daily_surcharge = 3.50
  ),
  open_plus = list(
    name = "Open Plus",
    hourly_rate = 7.65,
    first_daily_max = 50,
    additional_day = 36.50,
    km_interval_1 = 50,
    km_interval_rate_1 = 0.30,
    km_interval_rate_2 = 0.29,
    monthly_fee = 5,
    weekend_hourly_surcharge = 0.35,
    weekend_daily_surcharge = 3.50
  ),
  value = list(
    name = "Value",
    hourly_rate = 4.65,
    first_daily_max = 36.50,
    additional_day = 36.50,
    km_interval_1 = 50,
    km_interval_rate_1 = 0.51,
    km_interval_rate_2 = 0.38,
    monthly_fee = 5,
    weekend_hourly_surcharge = 0.35,
    weekend_daily_surcharge = 3.50
  ),
  value_plus = list(
    name = "Value Plus",
    hourly_rate = 4.05,
    first_daily_max = 30.50,
    additional_day = 30.50,
    km_interval_1 = 50,
    km_interval_rate_1 = 0.43,
    km_interval_rate_2 = 0.34,
    monthly_fee = 12.50,
    weekend_hourly_surcharge = 0.35,
    weekend_daily_surcharge = 3.50
  ),
  value_extra = list(
    name = "Value Extra",
    hourly_rate = 3.75,
    first_daily_max = 26.50,
    additional_day = 26.50,
    km_interval_1 = 0,
    km_interval_rate_1 = 0.34,
    km_interval_rate_2 = 0.34,
    monthly_fee = 30,
    weekend_hourly_surcharge = 0.35,
    weekend_daily_surcharge = 3.50
  ),
  value_extra_workday = list(
    name = "Value Extra (Workday)",
    hour_limit = 10,
    daily_rate = 25,
    km_interval_1 = 40,
    km_interval_rate_1 = 0,
    km_interval_rate_2 = 0.39,
    monthly_fee = 30
  ),
  flex = list(
    name = "Flex",
    minutely_rate = 0.43,
    hourly_max_ = 14.25,
    first_daily_max = 50,
    additional_day = 50,
    km_interval_1 = 75,
    km_interval_rate_1 = 0,
    km_interval_rate_2 = 0.31,
    weekend_hourly_surcharge = 0,
    weekend_daily_surcharge = 0
  )
)

long_distance_rates = list(
  low_season = list(
    first_day = 46,
    additional_days = 37,
    hourly_rate = 15,
    weekly_rate = 220,
    km_interval_1 = 300,
    km_interval_rate_1 = 0.29,
    km_interval_rate_2 = 0.19
  ),
  high_season = list(
    first_day = 61,
    additional_days = 52,
    hourly_rate = 15,
    weekly_rate = 290,
    km_interval_1 = 300,
    km_interval_rate_1 = 0.29,
    km_interval_rate_2 = 0.19
  )
)

time_cost_fn <- function(
  hours,
  plan
) {
  time_cost <- 0

  for (
    day in seq(
      0,
      hours %/% 24 + 0
    )
  ) {
    if (
      day == 0
    ) {
      time_cost <- time_cost + min(
        plan$hourly_rate*(
          hours - 24*day
        ),
        plan$first_daily_max
      )
    } else {
      time_cost <- time_cost + min(
        plan$hourly_rate*(
          hours - 24*day
        ),
        plan$additional_day
      )
    }
  }
  
  return(
    time_cost
  )
}

weekend_time_cost_fn <- function(
  hours,
  plan
) {
  weekend_time_cost <- 0

  for (
    day in seq(
      0,
      hours %/% 24 + 0
    )
  ) {
    if (
      day == 0
    ) {
      weekend_time_cost <- weekend_time_cost + min(
        (
          plan$hourly_rate + plan$weekend_hourly_surcharge
        )*(
          hours - 24*day
        ),
        plan$first_daily_max + plan$weekend_daily_surcharge
      )
    } else {
      weekend_time_cost <- weekend_time_cost + min(
        (
          plan$hourly_rate + plan$weekend_hourly_surcharge
        )*(
          hours - 24*day
        ),
        plan$additional_day + plan$weekend_daily_surcharge
      )
    }
  }
  
  return(
    weekend_time_cost
  )
}

distance_cost_fn <- function(
  distance,
  plan
) {
  distance_cost <- 0

  distance_cost <- plan$km_interval_rate_1*min(
    distance,
    plan$km_interval_1
  )

  if (
    distance > plan$km_interval_1
  ) {
    distance_cost <- distance_cost + plan$km_interval_rate_2*(
      distance - plan$km_interval_1
    )
  }

  return(
    distance_cost
  )
}

get_costs <- function(
  plan,
  hours,
  distance
) {
  data <- expand.grid(
    hours=hours,
    distance=distance
  )

  data <- data %>%
    mutate(
      time_cost = lapply(
        .$hours,
        time_cost_fn,
        plan = plan
      ) |> as.numeric(),
      weekend_time_cost = lapply(
        .$hours,
        weekend_time_cost_fn,
        plan = plan
      ) |> as.numeric(),
      distance_cost = lapply(
        .$distance,
        distance_cost_fn,
        plan = plan
      ) |> as.numeric(),
      total_cost = time_cost + distance_cost,
      weekend_total_cost= weekend_time_cost + distance_cost,
      plan_name = plan$name,
      tooltip = paste0(
        "$",
        total_cost,
        "\nTime (",
        hours,
        "h): $",
        time_cost,
        "\nDistance (",
        distance,
        "km): $",
        distance_cost
      )
    )

    return(data)
}

make_ggplotly <- function(
  data,
  max_cost
) {
  p <- ggplot(
    data,
    aes(
      hours,
      distance,
      fill=time_cost+distance_cost,
      text=tooltip
    )
  ) + 
    geom_tile() +
    scale_fill_gradient(
      low = "white", 
      high = "red",
      limits = c(
        0,
        max_cost
      )
    ) + 
    labs(
      fill = "Cost ($)",
      x = "Hours",
      y = "Distance (km)"
    ) +
    theme_minimal()

    return(
      ggplotly(
        p,
        tooltip="tooltip"
      )
    )
}
```

```{r "Compute some dfs - short_trips"}

hours <- seq(0,8,0.5)
distance <- seq(0,50,5)

costs_dfs = list()

for (
  plan_name in names(
    station_based_plans
  )
) {
  if (
    !(
      plan_name %in% c(
        "flex",
        "value_extra_workday"
      )
    )
  ) {
    costs_dfs[[plan_name]] <- get_costs(
      station_based_plans[[plan_name]],
      hours,
      distance
    )
  }
}

max_cost <- max(
  unlist(
    lapply(
      costs_dfs,
      \(x) max(
        x$total_cost
      )
    )
  )
)

best_cost_df <- costs_dfs |>
  bind_rows() |>
  group_by(
    hours,
    distance
  ) |>
  arrange(
    total_cost,
    plan_name,
    .by_group = TRUE
  ) |>
  mutate(
    tooltip = paste0(
      hours,
      " hours, ",
      distance,
      "km:\n---\n",
      plan_name,
      ": $",
      total_cost,
      " ($",
      time_cost,
      " + $",
      distance_cost,
      ")\n",
      collapse=""
    )
  ) |>
  slice_min(
    total_cost
  )
```

Fees not included:

- A minimum of 4 hours will be charged even for shorter trips if the final applied cost is that of the station-based vehicles of your plan
- Surcharge on displayed prices (time + km) of 15% for minivans and 10% for family vehicles.
- An additional charge of $10 per extra day applies to all trips longer than 28 days
- Damage protection plan
- Not considering FLEX pass
- Monthly membership fee
- Reducing Value plan to Open Plus

::: {.panel-tabset}

## Best plan

```{r "best plan heatmap"}
p <- ggplot(
  best_cost_df,
  aes(
    hours,
    distance,
    fill=plan_name,
    text=tooltip
  )
) + 
  geom_tile() +
  labs(
    fill = "Plan",
    x = "Hours",
    y = "Distance (km)"
  ) +
  scale_fill_paletteer_d(
    "nationalparkcolors::Acadia"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )

p

ggplotly(
  p,
  tooltip="tooltip"
)
```

## Open

```{r "open heatmap"}
make_ggplotly(
  costs_dfs$open,
  max_cost
)
```

## Open Plus

```{r "open plus heatmap"}
make_ggplotly(
  costs_dfs$open_plus,
  max_cost
)
```
## Value

```{r "value heatmap"}
make_ggplotly(
  costs_dfs$value,
  max_cost
)
```

## Value Plus

```{r "value plus heatmap"}
make_ggplotly(
  costs_dfs$value_plus,
  max_cost
)
```

## Value Extra

```{r "value extra heatmap"}
make_ggplotly(
  costs_dfs$value_extra,
  max_cost
)
```

:::